---
title: "final"
author: "Marcel, Owen"
date: "7/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Introduction

This markdown contains all analyses done for the manuscript "Contrasting resistance and resilience of the coupled oxic and anoxic components
of an experimental microbial community". The first part deals with the oxygen data, the second part deals with the microbial 16S rRNA amplicons (PacBio full length sequencing)

Short summary of the project:

-8 micro-ecosystems, incubated at 24 °C for 8 days under light:dark cycles of 16:8h
-afterwards, 4 micro-ecosystems (stressor treatment) were incubated in darkness for 7 days
-afterwards, the stressor-treated columns were incubated again as the controls 
-Oxygen was measured at the top and bottom of the liquid part continuously every 5 min
-samples were taken at day 8 (prior-stressor treatment), day 15 ( stressor-treatment), day 19 (short time recovery) and day 35 (long term recovery) at the height of the top and bottom oxygen sensor, respectively, for 16S rRNA full length amplicon sequencing

Aim of the study: Analysis of resistance and resilience of the oxygen concentration and the oxic/anoxic microbial community

# Preparation

```{r}
library(boot)
library(tidyverse)
library(lubridate)
library(googlesheets)
library(here)
library(readxl)
library(writexl)
library(ggplot2)
library(dplyr)
library(ggfortify)
library(ggpubr)
library(aod)
library(skimr)
library(scales)
library(patchwork)
library(broom)
library(aod)
library(Rcpp)
library(microbiome)
library(vegan)
library(plyr)
library(DESeq2)
library(apeglm)
library(patchwork)
```


# Oxygen 
## data handling

```{r}
oxygen_data <- read_excel("oxygen_data")

```

### Fix the date and times

```{r}
oxygen_data <- oxygen_data %>%
  mutate(Date_time = mdy_hms(paste0(Date, Time))) %>%
  select(-Date, -Time)
```


### Calculate hourly average

```{r}
hourly_value <- oxygen_data %>%
  mutate(Date = date(Date_time),
         Hour = hour(Date_time)) %>%
  group_by(Sensor_Name, Date, Hour) %>%
  dplyr::summarise(mean_Value = mean(Value)) %>%
  mutate(Date_time = ymd_h(paste0(Date, Hour)))

hourly_value <- hourly_value %>% mutate(Day = as.numeric(Date - ymd("2020-12-02")))
```

### Calculate the DAILY mean, minimum, and maximum of the *hourly* means
```{r}
daily_values <- hourly_value %>%
  ungroup() %>%
  group_by(Date, Sensor_Name) %>%
  dplyr::summarise(mean_O2 = mean(mean_Value),
            min_O2 = min(mean_Value),
            max_O2 = max(mean_Value),
            amplitude_O2 = max_O2 - min_O2) %>%
  mutate(Date_time_midpoint = ymd_hms(paste0(Date, "12:00:00"))) %>%
  pivot_longer(names_to = "Variable", values_to = "Oxygen", cols = 3:6)
```

### put in missing values for the days where there is no data
```{r}

missing_data <- crossing(Sensor_Name = unique(daily_values$Sensor_Name),
                         Variable = unique(daily_values$Variable),
                         Date = dmy(paste(c(24:27), "12-2020")),
                         Date_time_midpoint = ymd_hms(paste0(Date, " 12:00:00")))%>%
  mutate(Oxygen = NA)

daily_values <- rbind(daily_values, missing_data) %>%
  mutate(Sensor_Name=tolower(Sensor_Name)) %>%
  separate(Sensor_Name, into= c("Treatment", "Sensorposition","Column"), 
           remove=FALSE)

daily_values <- daily_values %>% mutate(Day = as.numeric(Date - ymd("2020-12-02")),
                                Variable2 = factor(Variable, levels = c("mean_O2", "max_O2", "min_O2", "amplitude_O2")))


```

## plot the data

## hourly values 
```{r}
plot_oxygen_a  <- hourly_value %>%
  mutate(Day2 = Day*24 + Hour) %>%
  ggplot(aes(x = Day2, y = mean_Value, col = Sensor_Name)) +
  geom_line()  +
  geom_rect(xmin = 195, xmax = 360,ymin = 0, ymax = 100, col = "lightblue", fill= "lightblue") +
  geom_line()  +
  geom_rect(xmin = 510, xmax = 640,ymin = 0, ymax = 100, col = "white", fill= "white") +
  scale_color_manual(values =c("Control_top_1" = "black", "Control_top_2" = "black", "Control_top_3" = "black", "Control_top_4" = "black","Disturbance_top_1" ="red","Disturbance_top_2" ="red","Disturbance_top_3" ="red","Disturbance_top_4" ="red","Control_bottom_1" = "grey", "Control_bottom_2" = "grey", "Control_bottom_3" = "grey", "Control_bottom_4" = "grey","Disturbance_bottom_1" ="grey","disturbance_bottom_2" ="grey","disturbance_bottom_3" ="grey","Disturbance_bottom_4" ="grey")) +
    theme_bw() + theme(panel.grid = element_blank()) + scale_x_continuous(breaks = seq(0, 817, by=48), labels=seq(0,35, by=2))+
  theme(legend.position = "none") +
  ylab("Hourly mean oxygen value [%]") +
  xlab("Day") +
  geom_point(data=data.frame(Day2=c(190, 360, 456, 840), mean_Value=rep(-1, 4)), colour="black", size=3, shape=4)


plot(plot_oxygen_a)

```

## just top sensor with all 4 variables and daily mean
```{r}

plot_oxygen_b<- daily_values %>%
  filter(Sensor_Name %in% c("control_top_1","control_top_2","control_top_3","control_top_4","disturbance_top_1","disturbance_top_2","disturbance_top_3","disturbance_top_4")) %>%
  ggplot(aes(x = Day, y = Oxygen, col = Sensor_Name)) +
  geom_rect(xmin = 8, xmax = 15,
          ymin = -Inf, ymax = 100, col = "lightblue", fill="lightblue", alpha=.1) +
  geom_line()  +
  facet_wrap( ~ Variable2, ncol = 4) +
  scale_color_manual(values =c("control_top_1" = "black", "control_top_2" = "black", "control_top_3" = "black", "control_top_4" = "black","disturbance_top_1" ="red","disturbance_top_2" ="red","disturbance_top_3" ="red","disturbance_top_4" ="red","control_bottom_1" = "black", "control_bottom_2" = "black", "control_bottom_3" = "black", "control_bottom_4" = "black","disturbance_bottom_1" ="red","disturbance_bottom_2" ="red","disturbance_bottom_3" ="red","disturbance_bottom_4" ="red")) +
  theme_bw() + theme(panel.grid = element_blank())+
  theme(legend.position = "none") +
  geom_point(data=data.frame(Day=c(7.5, 15, 19, 35), Oxygen=rep(-7, 4)), shape=4 ,colour="black", size=2) +
  ylab("Daily mean oxygen value [%]") +
  xlab("Day")

plot(plot_oxygen_b)
```


# Applying t-test for every day to get p.value for top sensors

```{r}

analysis <- daily_values %>%
  filter(Sensor_Name != "control_top_1") %>%
  na.omit() %>%
  group_by(Date, Variable, Date_time_midpoint, Sensorposition) %>%
  do(m1 = tidy(lm(Oxygen ~ Treatment, data=.), conf.int=T, conf.level = 0.95)) %>%
  unnest() %>%
  filter(term=="Treatmentdisturbance", Sensorposition == "top")

analysis <- analysis %>% mutate(Day = as.numeric(Date - ymd("2020-12-02")),
                                Variable2 = factor(Variable, levels = c("mean_O2", "max_O2", "min_O2", "amplitude_O2")))
```

## Plot the estimates

```{r}
plot_estimate_oxygen<- analysis %>%
  ggplot(aes(x = Day, y = estimate)) +
  geom_rect(xmin = 8, xmax = 15,
          ymin = -Inf, ymax = 100, col = "lightblue", fill="lightblue", alpha=.1) +
  geom_point()  +
    geom_ribbon(aes(ymin=conf.low,ymax=conf.high, x=Day), alpha=.4)+
    geom_rect(xmin = 22, xmax = 25,
          ymin = -Inf, ymax = 100, col = "white", fill="white", alpha=1) +
  facet_wrap( ~ Variable2, ncol = 4) +
  scale_color_manual(values =c("control_top_1" = "black", "control_top_2" = "black", "control_top_3" = "black", "control_top_4" = "black","disturbance_top_1" ="red","disturbance_top_2" ="red","disturbance_top_3" ="red","disturbance_top_4" ="red","control_bottom_1" = "black", "control_bottom_2" = "black", "control_bottom_3" = "black", "control_bottom_4" = "black","disturbance_bottom_1" ="red","disturbance_bottom_2" ="red","disturbance_bottom_3" ="red","disturbance_bottom_4" ="red")) +
  theme_bw() + theme(panel.grid = element_blank())+ 
  geom_hline(yintercept=0, linetype="dashed") +
  geom_point(data=data.frame(Day=c(7.5, 15, 19, 35), estimate=rep(-25, 4)), colour="black", size=2, shape=4) +
  ylab("Estimate") +
  xlab("Day")

print(plot_estimate_oxygen)

```

# Microbial community

## Load phyloseq data, remove chloroplast sequences, just include samples from Experiment part III

```{r}
ps <- readRDS("ps_oxic_anoxic.rds")
ps_new <-  subset_taxa(ps, Order != "Chloroplast")
taxa_names(ps_new) <- paste0("Seq", seq(ntaxa(ps)))
```

## Calculation of alpha diversity index
```{r}

diversity_test<- estimate_richness(ps_new, split=TRUE, measures=NULL)
diversity_test$ID <- rownames(diversity_test) 
names <- read_excel("barcodes_FKIII.xlsx")
names1 <- as.data.frame(names)
names1$ID <- gsub("--", "..", names1$...1)
diversity_analysis <- merge(diversity_test, names1, by=c("ID"))

stderr <- function(x, na.rm=FALSE) {
  if (na.rm) x <- na.omit(x)
  sqrt(var(x)/length(x))
}

div_means <- diversity_analysis %>%
  dplyr::filter(Position %in% c("bottom", "top")) %>%
    dplyr::group_by(Position, Day, Treatment) %>%
    dplyr::summarize(Shannon_mean = mean(Shannon),
                     Shannon_SE = stderr(Shannon))

div_raw <- diversity_analysis


  
plot_a_diversity_top <- ggplot() +
  geom_rect(data=div_raw, aes(xmin = 8, xmax = 15,
          ymin = -Inf, ymax = Inf), col = "lightblue", fill="lightblue", alpha=0.1) +
  geom_point(data = subset(div_raw, Position == "top"), aes(x=Day, y=Shannon, col=Treatment,shape=Position), size=3, alpha=.7)+
  scale_shape_manual(values=c(17,16))+
  geom_point(data=subset(div_means, Position == "top"), aes(x=Day, y=Shannon_mean, col=Treatment), shape="+", size = 10, position = "dodge", width = 0.25) +
  geom_line(data=subset(div_means, Position =="top"), aes(x=Day, y=Shannon_mean, col=Treatment), linetype="dashed") +
    scale_colour_manual(values = c("Control" = "black", "Disturbance" = "red"))+
  theme_bw() + theme(panel.grid = element_blank())+ 
  theme(legend.position = "none")  +
  scale_x_continuous(limits=c(0,36),breaks = c(0,8,15,19,35)) +
  ggtitle("top layer community") +
  theme(plot.title = element_text(size = 11))

print(plot_a_diversity_top)

plot_a_diversity_bottom <- ggplot() +
  geom_rect(data=div_raw, aes(xmin = 8, xmax = 15,
          ymin = -Inf, ymax = Inf), col = "lightblue", fill="lightblue", alpha=0.1) +
  geom_point(data = subset(div_raw, Position == "bottom"), aes(x=Day, y=Shannon, col=Treatment,shape=Position), size=3, alpha=.7)+
  scale_shape_manual(values=c(17,16))+
  geom_point(data=subset(div_means, Position == "bottom"), aes(x=Day, y=Shannon_mean, col=Treatment), shape="+", size = 10, position = "dodge", width = 0.25) +
  geom_line(data=subset(div_means, Position =="bottom"), aes(x=Day, y=Shannon_mean, col=Treatment), linetype="dashed") +
    scale_colour_manual(values = c("Control" = "black", "Disturbance" = "red"))+
  theme_bw() + theme(panel.grid = element_blank())+ 
  theme(legend.position = "none")  +
  scale_x_continuous(limits=c(0,36),breaks = c(0,8,15,19,35)) +
  ggtitle("bottom layer community")+
  theme(plot.title = element_text(size = 11))

print(plot_a_diversity_bottom)
```

## estimates of the alpha diversity analysis
```{r}

 analysis_shannon <- div_raw %>%
  group_by(Day, Position) %>%
  do(m1 = tidy(lm(Shannon ~ Treatment, data=.), conf.int=T, conf.level = 0.95)) %>%
  unnest(cols = c(m1)) %>%
  filter(term=="TreatmentDisturbance")

# plot the estimates
plot_estimates_shannon_top <- analysis_shannon%>%
  filter(Position=="top") %>%
  ggplot(aes(x = Day, y = estimate)) +
  geom_rect(xmin = 8, xmax = 15,
          ymin = -Inf, ymax = 100, col = "lightblue", fill="lightblue", alpha=.1) +
  geom_point()  +
    geom_errorbar(aes(ymin=conf.low,ymax=conf.high, x=Day),width=0.5, size=0.5, color="black")+
  theme_bw() + theme(panel.grid = element_blank())+ 
  geom_hline(yintercept=0, linetype="dashed") +
  ylab("Estimate") +
  xlab("Day")  +
  scale_x_continuous(limits=c(0,37),breaks = c(0, 8,15,19,35))
 

print(plot_estimates_shannon_top)

plot_estimates_shannon_bottom <- analysis_shannon%>%
  filter(Position=="bottom") %>%
  ggplot(aes(x = Day, y = estimate)) +
  geom_rect(xmin = 8, xmax = 15,
          ymin = -Inf, ymax = 100, col = "lightblue", fill="lightblue", alpha=.1) +
  geom_point()  +
    geom_errorbar(aes(ymin=conf.low,ymax=conf.high, x=Day),width=0.5, size=0.5, color="black")+
  theme_bw() + theme(panel.grid = element_blank())+ 
  geom_hline(yintercept=0, linetype="dashed") +
  ylab("Estimate") +
  xlab("Day")  +
  scale_x_continuous(limits=c(0,37),breaks = c(0, 8,15,19,35)) 

print(plot_estimates_shannon_bottom)



#Anova 
## top community
diversity_analysis_top<- diversity_analysis %>%
  filter(Position=="top")

lm_treatment <- lm(Shannon ~ Treatment*Day, data = diversity_analysis_top)
autoplot(lm_treatment)
anova(lm_treatment)
summary(lm_treatment)
 
##top sensor shannon index is affected "postponed", and it does not show resilience
 Anova_treatment <- lm(Shannon ~ Treatment*as.factor(Day), data = diversity_analysis_top)
autoplot(Anova_treatment)
anova(Anova_treatment)
summary(Anova_treatment)

AIC(lm_treatment, Anova_treatment)

#bottom community
diversity_analysis_bottom<- diversity_analysis %>%
  filter(Position=="bottom")

##bottom sensor shannon index is affected directly, and it does show resilience
lm_treatment <- lm(Shannon ~ Treatment*Day, data = diversity_analysis_bottom)
autoplot(lm_treatment)
anova(lm_treatment)
summary(lm_treatment)

#bottom sensor shannon index is affected directly, and it does show resilience
Anova_treatment <- lm(Shannon ~ Treatment*as.factor(Day), data = diversity_analysis_bottom)
autoplot(Anova_treatment)
anova(Anova_treatment)
summary(Anova_treatment)

AIC(lm_treatment, Anova_treatment)
```

# Calculation rel abundance, just sequences that appear > 0.1 %, making dataframes for each column position

```{r}
ps_rel <- transform_sample_counts(ps_new, function(x) x / sum(x) )

relab_threshold <- 0.001

ps_relab <- filter_taxa(ps_rel, function(x) !(sum(x < relab_threshold) == length(x)), TRUE)
ntaxa(ps_new)
ntaxa(ps_relab)

ps_relative <- transform_sample_counts(ps_relab, function(x)  x / sum(x))
df_rel <- psmelt(ps_relative)
df_rel_top <- df_rel %>%
  filter(Position == "top")
df_rel_bottom <- df_rel %>%
  filter(Position == "bottom")

```


# NMDS

## NMDS Plot

```{r}
mds_whole <- ps_relative@otu_table %>%
  as.data.frame() %>%
  metaMDS(., 
        distance = "bray",
        k = 2, ## number of dimensions to reduce to
        try = 200, ## number of random starts to try
        autotransform = FALSE ## best not to use
)
## 0.16


mds_whole_res <- ps_relative @sam_data %>%
  as.tibble() %>%
  select(Treatment, Column, Day, Position,Day_name) %>%
  bind_cols(as.tibble(scores(mds_whole, display = "sites")))

nmds_plot_supplements <- ggplot(mds_whole_res, aes(x = NMDS1, y = NMDS2)) +
  geom_point(aes(shape = Position, color=Treatment),
             size = 2) +
  scale_shape_manual(values =c(17,16))+
  facet_wrap("Day", ncol=2)+
   theme_bw() + theme(panel.grid = element_blank())+ 
  scale_colour_manual(values = c("Control" = "black", "Disturbance" = "red"))

print(nmds_plot_supplements)
```

# NMDS separated for NMDS1 and NMDS2 axis

```{r}
NMDS_top <- mds_whole_res %>% 
 filter(Position=="top") 

NMDS_bottom <- mds_whole_res %>% 
 filter(Position=="bottom") 


# Dynamics
plot_NMDS1_top <- NMDS_top %>%
  ggplot(aes(x = Day, y = NMDS1, col = Treatment)) +
  geom_rect(xmin = 8, xmax = 15,
          ymin = -Inf, ymax = 100, col = "lightblue", fill="lightblue", alpha=.1) +
  geom_point(shape=16, size = 2)+
  geom_smooth() +
  theme_bw() + theme(panel.grid = element_blank())+ 
    scale_colour_manual(values = c("Control" = "black", "Disturbance" = "red"))+
  ggtitle("top layer community") +
  scale_x_continuous(limits=c(0,36),breaks = c(0, 8,15,19,35))+
  theme(legend.position = "none")+
  theme(plot.title = element_text(size = 11))

print(plot_NMDS1_top)

plot_NMDS2_top <- NMDS_top %>%
  ggplot(aes(x = Day, y = NMDS2, col = Treatment)) +
  geom_rect(xmin = 8, xmax = 15,
          ymin = -Inf, ymax = 100, col = "lightblue", fill="lightblue", alpha=.1) +
  geom_point(shape=16, size = 2) +
  geom_smooth() +
  theme_bw() + theme(panel.grid = element_blank())+ 
  ggtitle("top layer community")+ 
  scale_x_continuous(limits=c(0,36),breaks = c(0, 8,15,19,35))+
    scale_colour_manual(values = c("Control" = "black", "Disturbance" = "red"))+
  theme(legend.position = "none")+
  theme(plot.title = element_text(size = 11))

print(plot_NMDS2_top)

plot_NMDS1_bottom <- NMDS_bottom %>%
  ggplot(aes(x = Day, y = NMDS1, col = Treatment)) +
  geom_rect(xmin = 8, xmax = 15,
          ymin = -Inf, ymax = 100, col = "lightblue", fill="lightblue", alpha=.1) +
  geom_point(shape=17, size = 2) +
  geom_smooth() +
  theme_bw() + theme(panel.grid = element_blank())+ 
    scale_colour_manual(values = c("Control" = "black", "Disturbance" = "red"))+
  ggtitle("bottom layer community")+ 
  scale_x_continuous(limits=c(0,36),breaks = c(0, 8,15,19,35))+
  theme(legend.position = "none")+
  theme(plot.title = element_text(size = 11))

print(plot_NMDS1_bottom)

plot_NMDS2_bottom<- NMDS_bottom %>%
  ggplot(aes(x = Day, y = NMDS2, col = Treatment)) +
  geom_rect(xmin = 8, xmax = 15,
          ymin = -Inf, ymax = 100, col = "lightblue", fill="lightblue", alpha=.1) +
  geom_point(shape=17, size = 2) +
  geom_smooth() +
  theme_bw() + theme(panel.grid = element_blank())+ 
  scale_x_continuous(limits=c(0,36),breaks = c(0, 8,15,19,35))+
    scale_colour_manual(values = c("Control" = "black", "Disturbance" = "red"))+
  ggtitle("bottom layer community")+ 
  theme(legend.position = "none")+
  theme(plot.title = element_text(size = 11))

print(plot_NMDS2_bottom)

```

## estimate analysis for NMDS data
```{r}

analysis_NMDS1_top <- NMDS_top %>%
  group_by(Day, Day_name) %>%
  do(m1 = tidy(lm(NMDS1 ~ Treatment, data=.), conf.int=T, conf.level = 0.95)) %>%
  unnest(cols = c(m1)) %>%
  filter(term=="TreatmentDisturbance")

plot_estimate_NMDS1_top <- analysis_NMDS1_top %>%
  ggplot(aes(x = Day, y = estimate)) +
  geom_rect(xmin = 8, xmax = 15,
          ymin = -Inf, ymax = 100, col = "lightblue", fill="lightblue", alpha=.1) +
  geom_point()  +
  geom_errorbar(aes(x=Day, ymin=conf.low, ymax=conf.high), width=0.5, size=0.5, color="black")+ 
  theme_bw() + theme(panel.grid = element_blank())+ 
  geom_hline(yintercept=0, linetype="dashed") +
  scale_x_continuous(limits=c(0,36),breaks = c(0, 8,15,19,35))+
  ylab("Estimate") +
  xlab("Day")

print(plot_estimate_NMDS1_top)


analysis_NMDS2_top <- NMDS_top %>%
  group_by(Day, Day_name) %>%
  do(m1 = tidy(lm(NMDS2 ~ Treatment, data=.), conf.int=T, conf.level = 0.95)) %>%
  unnest(cols = c(m1)) %>%
  filter(term=="TreatmentDisturbance")

plot_estimate_NMDS2_top <- analysis_NMDS2_top %>%
  ggplot(aes(x = Day, y = estimate)) +
  geom_rect(xmin = 8, xmax = 15,
          ymin = -Inf, ymax = 100, col = "lightblue", fill="lightblue", alpha=.1) +
  geom_point()  +
  geom_errorbar(aes(x=Day, ymin=conf.low, ymax=conf.high), width=0.5, size=0.5, color="black")+
  theme_bw() + theme(panel.grid = element_blank())+ 
  scale_x_continuous(limits=c(0,36),breaks = c(0, 8,15,19,35))+
  geom_hline(yintercept=0, linetype="dashed") +
  ylab("Estimate") +
  xlab("Day")

print(plot_estimate_NMDS2_top)

analysis_NMDS1_bottom <- NMDS_bottom %>%
  group_by(Day, Day_name) %>%
  do(m1 = tidy(lm(NMDS1 ~ Treatment, data=.), conf.int=T, conf.level = 0.95)) %>%
  unnest(cols = c(m1)) %>%
  filter(term=="TreatmentDisturbance")

plot_estimate_NMDS1_bottom <- analysis_NMDS1_bottom%>%
  ggplot(aes(x = Day, y = estimate)) +
  geom_rect(xmin = 8, xmax = 15,
          ymin = -Inf, ymax = 100, col = "lightblue", fill="lightblue", alpha=.1) +
  geom_point()  +
  geom_errorbar(aes(x=Day, ymin=conf.low, ymax=conf.high), width=0.5, size=0.5, color="black")+
  theme_bw() + theme(panel.grid = element_blank())+ 
  geom_hline(yintercept=0, linetype="dashed")+
  scale_x_continuous(limits=c(0,36),breaks = c(0, 8,15,19,35))+
  ylab("Estimate") +
  xlab("Day")

print(plot_estimate_NMDS1_bottom)

analysis_NMDS2_bottom <- NMDS_bottom %>%
  group_by(Day, Day_name) %>%
  do(m1 = tidy(lm(NMDS2 ~ Treatment, data=.), conf.int=T, conf.level = 0.95)) %>%
  unnest(cols = c(m1)) %>%
  filter(term=="TreatmentDisturbance")

plot_estimate_NMDS2_bottom <- analysis_NMDS2_bottom%>%
  ggplot(aes(x = Day, y = estimate)) +
  geom_rect(xmin = 8, xmax = 15,
          ymin = -Inf, ymax = 100, col = "lightblue", fill="lightblue", alpha=.1) +
  geom_point()  +
  geom_errorbar(aes(x=Day, ymin=conf.low, ymax=conf.high), width=0.5, size=0.5, color="black")+
  theme_bw() + theme(panel.grid = element_blank())+ 
  geom_hline(yintercept=0, linetype="dashed") +
  scale_x_continuous(limits=c(0,36),breaks = c(0, 8,15,19,35))+
  ylab("Estimate") +
  xlab("Day")

print(plot_estimate_NMDS2_bottom)
```

# final plot of figures combined

```{r}
p1 <- plot_oxygen_a
p2 <- plot_oxygen_b
p3 <- plot_a_diversity_top
p4<-  plot_a_diversity_bottom
p5 <- plot_estimates_shannon_top
p6 <- plot_estimates_shannon_bottom
p7 <- plot_NMDS1_bottom
p8 <- plot_NMDS2_bottom
p9 <- plot_NMDS1_top
p10 <- plot_NMDS2_top

p11 <- plot_estimate_oxygen
p12<- plot_estimates_shannon_top
p13<- plot_estimates_shannon_bottom
p14 <- plot_estimate_NMDS1_bottom
p15 <- plot_estimate_NMDS2_bottom
p16 <- plot_estimate_NMDS1_top
p17 <- plot_estimate_NMDS2_top


#Figure 1
plot_1 <- p1/
  p2/
  p11 

plot_1 +
  plot_layout(heights=c(2,1,1)) + plot_annotation(tag_levels = "a")


#Figure 2
plot <- (p4|p13)/
  (p3|p12)/
  (p7|p14)/
  (p8|p15)/
  (p9 |p16)/
  (p10 |p17)
  

plot + plot_annotation(tag_levels = "a")
```

#Supplementary Information

## Visualization of rel.abundance for supplementary information
```{r}
df_class<- df_rel %>%
  group_by(Class, Treatment,Day, Column, Position) %>%
  dplyr::summarize(Abundance=sum(Abundance))


index <- which(df_class$Abundance>=0.05)
class_to_keep <- unique(df_class[index,"Class"])
class_to_keep <- unname(unlist(class_to_keep))

 

df_class$Class_filter <- ifelse(df_class$Class %in% class_to_keep, df_class$Class,"Zother")

   cols <-c("Actinobacteria"="bisque1", "Alphaproteobacteria" = "yellow", "Babeliae"="lightcyan1","Bacilli" = "green4",  "Bacteroidia" ="maroon1", "Campylobacteria" ="royalblue1","Zother"="black","Chlorobia" ="thistle1", "Clostridia"= "firebrick1", "Cyanobacteriia" = "magenta4", "Fimbriimonadia" = "grey","Gammaproteobacteria"="red", "Microgenomatia"="grey38","Oligoflexia"="snow1", "Phycisphaerae"="cornflowerblue", "Planctomycetes"="lightskyblue1","Spirochaetia"="lavender", "Vampirivibronia" = "lightsalmon", "Verrucomicrobiae" = "slateblue3")
   
  supplement_plot1 <-df_class%>%
  ggplot(aes_string(x = "Column", y = "Abundance", fill="Class_filter" )) +
     geom_bar(stat = "identity", position = "stack", col="black") +
       facet_wrap(Position~Day, ncol = 4)  +
       theme(axis.text.x = element_text(angle = -45, hjust = 0, vjust=0.5, size=9))+
    ylab("Relative abundance") +
  scale_fill_manual("legend", values = cols)
  
  print(supplement_plot1)
```

